---
title: "Reproducible Research: Peer Assessment 1"
date: "2/24/2018"
output: html_document
keep_md: true
---

## Loading and preprocessing the data
```{r echo=FALSE}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = paste0(getwd(), '/repdata%2Fdata%2Factivity.zip'), method = "curl")
unzip("repdata%2Fdata%2Factivity.zip",exdir = "data")
activity <- read.csv("activity.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```
Formatting the data column
```{r echo=TRUE}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
str(activity)
```
Check the dimensions
```{r echo=TRUE}
dim(activity)
```
```{r echo=TRUE}
head(activity)
```

## 1. What is mean total number of steps taken per day?
Group and summarize the data using dplyr
```{r echo=TRUE}
library(dplyr)
AvgDay <- activity %>% group_by(date) %>%
          summarize(total.steps = sum(steps, na.rm = T), 
                  mean.steps = mean(steps, na.rm = T))
```
Make that histogram with ggplot!
```{r echo=TRUE}
library(ggplot2)
g <- ggplot(AvgDay, aes(x=total.steps))
g + geom_histogram(binwidth = 2000) + theme(axis.text = element_text(size = 10),  
      axis.title = element_text(size = 10)) + labs(y = "Frequency") + labs(x = "Total steps per day")
```

```{r echo=TRUE}
summary(AvgDay$total.steps)
```

## 2. What is the daily activity pattern?
Make the time series plot
```{r echo=TRUE}
AvgInterval <- activity %>% group_by(interval) %>%
      summarize(mean.steps = mean(steps, na.rm = T))
g <- ggplot(AvgInterval, aes(x = interval, y = mean.steps))
g + geom_line() + theme(axis.text = element_text(size = 10), 
      axis.title = element_text(size = 10, face = "bold")) + 
      labs(y = "Mean number steps") + labs(x = "Interval")
```
So we can see the the maximum number of steps is 206 and that occurred at time interval 835.

## 3. Imputing mission values
How many NAs do we have? And where are they?
```{r echo=TRUE}
table(is.na(activity) == TRUE)
```
```{r echo=TRUE}
sum(is.na(activity$steps))
```
```{r echo=TRUE}
summary(activity)
```
We have 2304 NAs, that are all in the steps category

### Strategy for filling in the missing values
Replace the NAs with the mean, 61
```{r echo=TRUE}
activity2 <- activity
for (i in 1:nrow(activity)){
    if(is.na(activity$steps[i])){
        index <- activity2$interval[i]
	value <- subset(AvgInterval, interval==index)
	activity2$steps[i] <- value$mean.steps
    }
}
```
```{r echo=TRUE}
newAvg <- activity2%>% group_by(date) %>%
      summarize(total.steps = sum(steps, na.rm = T))
```

Now the missing values are filled up, so let's create the histogram
```{r echo=TRUE}
g <- ggplot(newAvg, aes(x=total.steps))
g + geom_histogram(binwidth = 2500) + theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 14)) + labs(y = "Frequency") + labs(x = "Total steps/day")
```
What is the new median and mean?
```{r echo=TRUE}
summary(newAvg$total.steps)
```

## 4. Difference between the weekdays and weekends?
```{r echo=TRUE}
activity2$day <- ifelse(weekdays(activity2$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
```
```{r echo=TRUE}
wkend <- filter(activity2, day == "weekend")
wkday <- filter(activity2, day == "weekday")
```
```{r echo=TRUE}
wkend <- wkend %>%
      group_by(interval) %>%
      summarize(mean.steps = mean(steps)) 
wkend$day <- "weekend"

wkday <- wkday %>%
      group_by(interval) %>%
      summarize(mean.steps = mean(steps)) 
wkday$day <- "weekday"

newInterval <- rbind(wkend, wkday)
newInterval$day <- as.factor(newInterval$day)
newInterval$day <- relevel(newInterval$day, "weekend")
```
```{r echo=TRUE}
g <- ggplot (newInterval, aes (interval, mean.steps))
g + geom_line() + facet_grid (day~.) + theme(axis.text = element_text(size = 12), 
      axis.title = element_text(size = 14)) + labs(y = "Number of Steps") + labs(x = "Interval")
```
cool